Six decades of deforestation and forest fragmentation in Madagascar

### 2017-02-06 deforestmap.Rmd
Add tables

---
title: "Sixty years of deforestation and forest degradation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

@@

Accurate maps of forest cover and forest cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016]]. Unfortunately, accurate and cloud-free forest cover maps are not available for Madagascar for the last decade (2005-2015). In their paper untitled _"Fifty years of deforestation and forest degradation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000). The 1950s forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent for Madagascar in the 1950s. Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 30m and can be used to derive much more accurate estimates of the forest cover at these three dates. Notheless, maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example). Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013]. Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015]. However, these maps might (i) be too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests in @ONE2015). Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm. In addition, visual interpretation might be subjective and is potentially a source of errors. All this work of image selection and visual interpretation will have to be repeated to produce new forest maps in the future.

Global forest products have also been published recently and could be used to derive recent forest cover map at the national scale for Madagascar. @Kim2015 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000. Also, the approach used in @Kim2015 did not allow to correctly map the forest in the dry and spiny ecosystems. Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012. Meanwhile, results have been updated and are available up to the year 2014 at http://earthenginepartners.appspot.com/science-2013-global-forest. One problem with using the @Hansen2013 product to map forest cover is that a tree cover threshold must be defined to transform the tree cover data into a forest/non-forest map. Identifying such threshold is not straightforward as forests cannot be defined only by a tree cover value. Also, the tree cover threshold might strongly vary between forest types, mean tree cover values being substantially lower for dry forests than for moist forests. Also, a tree cover gives no information on land-use. Therefore, tree plantations such as oil palm or eucalyptus plantations cannot be differenciated from natural forests, which provide very different ecosystem services particularly in terms of biodiversity [@Tropek2014]. Thus, using @Hansen2013 alone can thus only give a first-approach estimate of the forest cover. But, tree cover is not the only information provided by @Hansen2013. The forest loss, i.e. the complete loss of forest tree cover (which shifts from a positive value over 10% to zero) is also reported.

@@

### 3.2 Evolution of forest cover, deforestation intensity and fragmentation with time

@@

### 4.1 Evolution of deforestation and forest fragmentation on the last 60 years in Madagascar

@@

>>>>

---
title: "Six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

@@

Accurate maps of forest cover and forest cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016]]. Unfortunately, accurate and cloud-free forest cover maps are not available for Madagascar for the last decade (2005-2015). In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000). The 1950s forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s. Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 30m and can be used to derive much more accurate estimates of the forest cover at these three dates. Notheless, maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example). Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013]. Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015]. However, these maps might (i) be too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests in @ONE2015). Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm. In addition, visual interpretation might be subjective and is potentially a source of errors. All this work of image selection and visual interpretation will have to be repeated to produce new forest maps in the future.

Global forest products have also been published recently and could be used to derive recent forest cover map at the national scale for Madagascar. @Kim2014 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000. Also, the approach used in @Kim2014 did not allow to correctly map the forest in the dry and spiny ecosystems. Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012. Meanwhile, results have been updated and are available up to the year 2014 at http://earthenginepartners.appspot.com/science-2013-global-forest. One problem with using the @Hansen2013 product to map forest cover is that a tree cover threshold must be defined to transform the tree cover data into a forest/non-forest map. Identifying such threshold is not straightforward as forests cannot be defined only by a tree cover value. Also, the tree cover threshold might strongly vary between forest types, mean tree cover values being substantially lower for dry forests than for moist forests. Also, a tree cover gives no information on land-use. Therefore, tree plantations such as oil palm or eucalyptus plantations cannot be differenciated from natural forests, which provide very different ecosystem services particularly in terms of biodiversity [@Tropek2014]. Thus, using @Hansen2013 alone can thus only give a first-approach estimate of the forest cover. But, tree cover is not the only information provided by @Hansen2013. The forest loss, i.e. the complete loss of forest tree cover (which shifts from a positive value over 10% to zero) is also reported.

@@

Difference might be due to post-processing of @Rakotomalala2015 map.

### 3.2 Evolution of forest cover and deforestation intensity

### 3.3 Evolution of forest fragmentation with time

@@

### 4.1 Evolution of deforestation and forest fragmentation in Madagascar on the last 60 years

@@

```{r comp, results="asis"}
tab.1 <- comp_moist.df
mean.diff <- round(100*(sum(tab.1$BSM)-sum(tab.1$PERR_FH))/sum(tab.1$PERR_FH))
names(tab.1) <- c("Year","Vieil (ha)", "Rakot (ha)", "Rakot_NA (ha)", "ad_Vieil (ha/yr)", "ad_Rakot (ha/yr)", "theta_Vieil (%/yr)", "theta_Rakoto (%/yr)", "diff.area (%)")
tab.1 <- tab.1[,c(1:4,9,5:8)]
kable(tab.1)
```

Table 1: **Comparison of the forest cover and deforestation rates between our study and the study by @Rakotomalala2015**. This comparison is done only for the humid ecoregion for the periods 2005-2010-2013, which is the extent and time-periods of the @Rakotomalala2015 map. @Rakotomalala2015 map has been validated using 10,854 independant visual interpretation points leading to a rather high accuracy (Kappa=0.79 and overall-accuracy=0.89). Our study (suffix _"Vieil"_) estimated a forest cover greater by about `r mean.diff`% compared to @Rakotomalala2015 (suffix _"Rakot"_). But the difference (_"diff.area"_) did not vary much between dates (between 3.56 and 6.18%). Our study also estimated higher annual deforestation (_"ad"_) and deforestation rates (_"theta"_) but the same trend for the deforestation was observed between the two studies, with a doubling of the deforestation for the period 2010-2013 compared to 2005-2010. The differences between the two studies could be attributed to the forest definition which might be different in the two studies and to the post-processing in @Rakotomalala2015 map.

```{r defor, results="asis"}
tab.2 <- defor.df[,-c(2)]
kable(tab.2)
```

Table 2: **Evolution of the forest cover and deforestation intensity from 1950s to 2014 in Madagascar**.
